\documentclass{ximera}

\input{../preamble}


\title{Linear Dependence and Independence}
\author{\phantom{Dr. Golubitsky}}
\date{Due: TBA}


\begin{document}
\begin{abstract}
Linear Dependence and Independence.  
\end{abstract}
\maketitle



\problemlabel



\exerciselabel{2}{5.4}\begin{exercise} \label{c5.4.2}
For which values of $b$ are the vectors $(1,b)$ and $(3,-1)$
linearly independent?

\begin{solution}

\ans The set is linearly independent if $b \neq -\frac{1}{3}$.

\soln Note that a set of two vectors is linearly dependent if one is
a multiple of the other.  So this set is dependent for any values of
$b$ at which
\[
(3,-1) = \alpha(1,b).
\]
When equality holds $\alpha = 3$.  Therefore, $b = -\frac{1}{3}$.  

\end{solution}
\end{exercise}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\problemlabel

\exerciselabel{3}{5.4}\begin{exercise} \label{c5.4.3}
Let
\[
u_1=(1,-1,1) \quad u_2=(2,1,-2) \quad u_3 = (10,2,-6).
\]
Is the set $\{u_1,u_2,u_3\}$ linearly dependent or linearly independent?

\begin{solution}

\ans The set is linearly dependent.

\soln Let $A$ be the matrix whose columns are $u_1$, $u_2$, and $u_3$. 
The set $\{u_1,u_2,u_3\}$ is linearly dependent if there exists
a nonzero vector $r = (r_1,r_2,r_3)$ such that $r_1u_1 + r_2u_2 +
r_3u_3 = 0$, that is, if the homogeneous system $Ar = 0$ has a
nonzero solution.  Row reduce:
\[
\matthree{1}{2}{10}{-1}{1}{2}{1}{-2}{-6} \longrightarrow
\matthree{1}{0}{2}{0}{1}{4}{0}{0}{0}.
\]
So, $Ar = 0$ when $r = r_3(-2,-4,1)$.
The value of $r$ is nonzero for $r_3 \neq 0$, so the set is indeed
linearly dependent.
As an example, let $r_3 = 1$.  Then,
\[
-4u_1 - 2u_2 + u_3 = -2(1,-1,1) - 4(2,1,-2) + (10,2,-6) =
(0,0,0) = 0.
\]

\end{solution}
\end{exercise}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\problemlabel

\exerciselabel{5}{5.4}\begin{exercise}\label{C5.4.10}
 Let 
 \[
 u = \Matrix{3\\2\\-5} \quad v = \Matrix{-6\\1\\7} \quad w = \Matrix{0\\5\\-3}
 \]
\begin{enumeratea}
\item Determine whether the sets $\{u,v\}, \{u,w\}, \{v,w\}$ are linearly independent.
\item Is the set $\{u,v,w\}$ linearly independent? 
\end{enumeratea}

  
\begin{solution}

\ans %THE SHORT ANSWER GOES HERE (if relevant)
\begin{enumeratea}
\item yes, yes, yes.
\item no
\end{enumeratea}

\soln %A LONGER SOLUTION, EXPLAINING HOW THE ANSWER WAS FOUND, GOES HERE (if relevant)

\begin{enumeratea}
\item A set of two vectors is linearly dependent if and only if one is a scalar multiple of the other.  None of the vectors is a scalar multiple of another. Thus, $\{u,v\}, \{u,w\}, \{v,w\}$ are linearly independent.
\item To tell whether $\{u,v,w\}$ is linearly independent, we solve the equation $au + bv + cw = 0$:
\[ 
a \Matrix{3 \\ 2 \\ -5} + b\Matrix{-6\\1\\7} + c\Matrix{0\\5\\-3} = \Matrix{0\\0\\0}
\]
We row reduce the coefficient matrix of this system: 
\[ 
\Matrix{ 3 & -6 & 0 \\ 2 & 1 & 5 \\ -5 & 7 & -3 } \to \Matrix{ 1 & -2 & 0 \\ 2 & 1 & 5 \\ -5 & 7 & -3 }  \to \Matrix{ 1 & -2 & 0 \\ 0 & 5 & 5 \\ 0 & -3 & -3 } \to\Matrix{ 1 & -2 & 0 \\ 0 & 1 & 1 \\ -5 & 1 & -1 } \] \[ \to \Matrix{ 1 & -2 & 0 \\ 0 & 1 & 1 \\ 0 & 0 & 0 } \to  \Matrix{ 1 & 0 & 2 \\ 0 & 1 & 1 \\ 0 & 0 & 0 } 
\]
Therefore, there are infinitely many solutions to this linearly system and $\{u,v,w\}$ is linearly dependent.

\end{enumeratea}



\end{solution}
\end{exercise}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%



\end{document}
